Lens Modeling in \sw

### 2014-06-26 text/ms.tex

\usepackage{graphicx}
% prefere pdf over png (better quality, lossless compression)
\DeclareGraphicsExtensions{%
  .pdf,.PDF,%
  .png,.PNG,%
  .jpg,.mps,.jpeg,.jbig2,.jb2,.JPG,.JPEG,.JBIG2,.JB2}

@@

\author[Kueng et al]{Rafael Kueng,$^{1}$
Phil Marshall,$^{2}$
Anupreeta More,$^{3}$
Aprajita Verma,$^{4}$
Amit Kapadia,$^{5}$
\newauthor
Elisabeth Baeten,$^{6}$
Claude Cornen,$^{6}$
Christine McMillan,$^{6}$
Julianne Wilcox,$^{6}$
\newauthor
Jonas Odermatt,$^{7}$
Jonathan Coles,$^{8}$
Surhud More$^{3}$
and Prasenjit Saha$^{1}$ \\
$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
$^{2}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
$^{3}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
$^{4}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
$^{5}$Adler Planetarium, 1300 S. Lake Shore Drive, Chicago, IL 60605, USA\\
$^{6}$Spacewarps\\
$^{7}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
$^{8}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France}

>>>>

\usepackage{graphicx}
% prefer pdf over png (better quality, lossless compression)
\DeclareGraphicsExtensions{%
  .pdf,.PDF,%
  .png,.PNG,%
  .jpg,.mps,.jpeg,.jbig2,.jb2,.JPG,.JPEG,.JBIG2,.JB2}

@@

\author[Kueng et al]{Rafael Kueng,$^{1}$
Phil Marshall,$^{2}$
Anupreeta More,$^{3}$
Aprajita Verma,$^{4}$
\newauthor
Elisabeth Baeten,$^{5}$
Claude Cornen,$^{5}$
Christine McMillan,$^{5}$
Julianne Wilcox,$^{5}$
\newauthor
Jonas Odermatt,$^{6}$
Jonathan Coles,$^{7}$
Surhud More$^{3}$
and Prasenjit Saha$^{1}$ \\
$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
$^{2}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
$^{3}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
$^{4}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
$^{5}$Spacewarps\\
$^{6}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
$^{7}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France}
%$^{8}$Adler Planetarium, 1300 S. Lake Shore Drive, Chicago, IL 60605, USA\\

### 2014-06-26 text/partA.tex

The encouraging early results prompted the question: could modeling of
the lenses also be done by volunteers?  There are several software
tools for lensing modelling available, and work has been done on
generic interfaces \citep{2013NewAR..57....1L,2014A&C.....5...28L}.
Some early designs for \sw included a prototype lens modelling tool
\citep{2010AAS...21543527N}.  Moreover, some {\em Spacewarps\/}
volunteers are quite experienced from earlier projects, having spent a
thousand hours or more with data, and are very interested in more
demanding projects.  The interests of citizen-science communities are
just beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but
it is clear that some volunteers welcome open-ended challenges, and
sometimes these have led to new scientific results: one example is the
discovery of an exceptional extra-solar planet
\citep{2013ApJ...768..127S}; another is the development of new
algorithms for protein folding \citep{Khatib22112011}.  All these are
grounds for optimism.  There is, however, a basic difficulty in strong
gravitational lensing.  Lensed images do not look much like their
source, and still less do they resemble the lensing-mass distribution.
To model a lens, one needs to either do a lot of random guessing, or
have a good intuition for what works.

@@

\subsection{Diagnostics}

@@

The information in this section was not revealed to the main developer
of \spl (RK, who also chose the challenge set) or to the modellers
(EB, CC, CM, JO, PSand JW) until the modelling stage was done.

@@

Of the 119 models proposed, we discuss now present eight examples in
some detail.  Results from these are shown in Figures
\ref{fig:6941}--\ref{fig:7022}.  The first four of these show the most
common image morphologies, the other four explain some problem cases.

@@

@@

Figure \ref{fig:6941} shows a simple example.  In summary: the
identification of minimum and saddle point is correct, but the
estimated Einstein radius is a little too high.

>>>>

The encouraging early results prompted the question: could modelling
of the lenses also be done by volunteers?  If so, modelling could help
assess prioritise lens candidates at an early stage, which would be
very useful with new wide-field and sensitive surveys, which will
yield thousands of lens candidates. There are several software tools
for lensing modelling available, and work has been done on generic
interfaces \citep{2014A&C.....5...28L}.  Some early designs for \sw
included a prototype lens modelling tool \citep{2010AAS...21543527N}.
Moreover, some {\em Spacewarps\/} volunteers are quite experienced
from earlier projects, having spent a thousand hours or more with
data, and are very interested in more demanding projects.  The
interests of citizen-science communities are just beginning to be
studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that
some volunteers welcome open-ended challenges, and sometimes these
have led to new scientific results: one example is the discovery of an
exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
the development of new algorithms for protein folding
\citep{Khatib22112011}.  All these are grounds for optimism.  There
is, however, a basic difficulty in strong gravitational lensing.
Lensed images do not look much like their source, and still less do
they resemble the lensing-mass distribution.  To model a lens, one
needs to either do a lot of random guessing, or have a good intuition
for what works.

@@

\subsection{Diagnostics} \label{sec:diag}

@@

The information in this section was not revealed to the main developer
of \spl (RK, who also chose the challenge set) or to the modellers
(EB, CC, CM, JO, PS and JW) while modelling was in progress.  After
the modelling stage, AM released the details of the sims for
post-modelling analysis.  Results from the latter now follow.

@@

Of the 119 models proposed, we now present eight examples in some
detail.  Results from these are shown in Figures
\ref{fig:6941}--\ref{fig:7022}.  The first four of these show the most
common image morphologies, the other four explain some problem cases.

@@

We should remark that the scales and alignments in the six panels are
not completely consistent: the three panels on the right exactly the
same coordinates.  The middle-left and lower-left panels are
consistent with each other, but only approximately aligned with the
right column, becuse the latter depends on inexact mouse placements.
The upper-left panel is zoomed out slightly, to accommodate all the
markup.  For qualitative features, which is all that
Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, these small
inconsistencies are harmless.

@@

Figure \ref{fig:6941} shows the simplest case, with two clear images
produced by a nearly-circular lens.  The centre of the lensing galaxy
is a maximum, the image nearer to the galaxy is a saddle point, and
image further away is a minimum, and these were correctly identified.
The model synthetic images is very crude, because there is no fitting
for the source light distribution, as already noted in \secref{diag}.
